---
title: "R Notebook"
output: html_notebook
---

# Forecasting the Construction of Residential Housing in the USA.

# Business Understading
We intend to learn to estimate future trends in the American housing market for small and
medium-sized dwellings based on previous patterns and present market circumstances. We also
seek to improve our knowledge of the variables influencing small and medium-sized housing
demand and supply, such as interest rates, economic growth, demography, and government
regulations. This forecasting will also aid in the identification of potential risks and uncertainties
in the housing market, such as price changes, housing oversupply or undersupply, and altering
consumer preferences.


# Data

Citations
Data Source Link –
https://www.census.gov/econ/currentdata/dbsearch?programCode=RESCONST&startYear=195
9&endYear=2023&categories[]=APERMITS&dataType=TOTAL&geoLevel=US&adjusted=1&
notAdjusted=0&errorData=0#table-results


## Import libraries
```{r}
library(tidymodels)
library(readr)
library(zoo)
library(lubridate)
library(forecast)
```


## Import dataset
```{r}
library(readr)
data <- read_csv("C:/Users/OMBATI/Desktop/R codes/timeseries/SeriesReport-202305021054.csv")
View(data)
head(data)
```

I will remove the first 5 rows and split the column into two;Period and Value

## Cleaning data;Removing the first 5 rows and assigning the 6th as the column name
```{r}
clean_data <- data[-c(1:5), ]

# extract column names from the first row
colnames(clean_data) <- clean_data[1,]

# remove the first row
clean_data <- clean_data[-1, ]

head(clean_data)

```


## Cleaning data;split the column into two;Period and Value
```{r}
# split the column into "Period" and "Value" columns
clean_data <- data.frame(do.call("rbind", strsplit(as.character(clean_data$"Period,Value"), ",")))
colnames(clean_data) <- c("Period", "Value")

# convert "Value" column to numeric
clean_data$Value <- as.numeric(clean_data$Value)

head(clean_data)

```
## EDA
```{r}
skimr::skim(clean_data)
```

```{r}
# convert the "Period" column to a date format
clean_data$Period <- as.Date(paste0("01-", clean_data$Period), format = "%d-%b-%Y")

# convert the data frame to a time series object
clean_data_ts <- ts(clean_data$Value, start = c(year(clean_data$Period[1]), month(clean_data$Period[1])),
                    frequency = 12)

# print the time series object
clean_data_ts
```

## Dealing with missing data in the time series data type
```{r}
# Check for missing values
sum(is.na(clean_data_ts))

# Impute missing values using na.interp function from forecast package
clean_data_ts_imputed <- na.interp(clean_data_ts)

# Check for missing values again
sum(is.na(clean_data_ts_imputed))
```


# Holt-Winter method for forecasting
```{r}
# apply Holt-Winter method for forecasting
hw_model <- hw(clean_data_ts)
hw_forecast <- forecast(hw_model, h = 12)
hw_forecast
```


# plot forecasts and actual values for Holt-Winters model
```{r}
autoplot(hw_model, include = 60) + 
  autolayer(clean_data_ts, series = "Actual values") + 
  xlab("Year") + ylab("Value") + ggtitle("Holt-Winters Forecast")
```


# Perform ETS method
```{r}
# Perform ETS method
ets_model <- ets(clean_data_ts)
ets_forecast <- forecast(ets_model, h = 12)
ets_forecast
```
# plot forecasts and actual values for ETS model
```{r}
autoplot(ets_model, include = 60) + 
  autolayer(clean_data_ts, series = "Actual values") + 
  xlab("Year") + ylab("Value") + ggtitle("ETS Forecast")
```



## Perform ARIMA method
```{r}
# Perform ARIMA method
arima_model <- auto.arima(clean_data_ts)
arima_forecast <- forecast(arima_model, h = 12)
arima_forecast
```
# plot forecasts and actual values for ARIMA model
```{r}
autoplot(arima_model, include = 60) + 
  autolayer(clean_data_ts, series = "Actual values") + 
  xlab("Year") + ylab("Value") + ggtitle("ARIMA Forecast")
```





